## Exercise (A) {.unnumbered}

::: callout-tip
You cannot input any codes on this website but need to download the exercise files on your computer. You can download all of the related files in a zip file **accessingEx.zip** from [Github folder](https://github.com/ehsanx/EpiMethods/tree/main/LabExercises/), or just by clicking this link [directly](https://github.com/ehsanx/EpiMethods/raw/main/LabExercises/accessingEx.zip).

-   Navigate to the GitHub folder (above link) where the ZIP file is located.
-   Click on the file name (above zip file) to open its preview window.
-   Click on the Download button to download the file. If you can't see the Download button, click on "Download Raw File" link that should appear on the page.
:::

## Problem Statement

We will use the article by @palis2020relationship, [DOI: 10.1080/09638237.2018.1437602](https://www.researchgate.net/profile/Heather_Palis/publication/323136807_The_relationship_between_sense_of_community_belonging_and_self-rated_mental_health_among_Canadians_with_mental_or_substance_use_disorders/links/5a96ff210f7e9ba42974c0dd/The-relationship-between-sense-of-community-belonging-and-self-rated-mental-health-among-Canadians-with-mental-or-substance-use-disorders.pdf).

-   Download the [CCHS MH topical index](https://datalib.usask.ca/dli/health/cchs/2012-mh/doc/cchs_mh_topical_index.pdf)
-   Download the [CCHS MH Data Dictionary](https://datalib.usask.ca/dli/health/cchs/2012-mh/doc/cchs_mh_datadictionary.pdf)

## Question 1: \[60% grade\]

### 1(a) Importing dataset

```{r}
# Importing dataset
load("Data/accessing/cchsMH.RData")
```

### 1(b) Subsetting according to eligibility

Subset the dataset according to the eligibility criteria / restriction specified in the paper

-   Identify the variable needed for eligibility criteria

**Hint**

-   Read the first paragraph of **Analytic sample** (page 2) for the eligibility criteria
-   Eligibility criteria was determined based on only **one** variable. Only work with 'YES' category.

a mental or substance use disorder (MSUD).

at least
one of the following MSUDs in the previous 12 months: 
a mood disorder (i.e. a major depressive episode, bipolar disorder, or mania), 
generalized anxiety disorder, or substance abuse or dependence (i.e. alcohol, cannabis, or other drugs).
Concurrent mental and substance use disorders could not be determined; 
however, just whether respondents reported the presence of any of these disorders at any point in the previous year.

```{r}
# your code here
colnames(cmh) %>% sort
table(cmh$ADM_RNO)
cmh_sub <- dplyr::filter(cmh, CCC_280=="YES"   )

```

### 1(c) Retaining necessary variables

In the dataset, retain only the variables associated with outcome measure, explanatory variable, potential confounders and survey weight. There should be eight variables (one outcome, one exposure, five confounders, and one survey weight).

Here are the steps:

-   Identify the outcome variable

-   Identify the explanatory variable

-   Identify the potential confounders

-   Identify the survey weight variable

-   **Hint**

    1.  Read

    -   first and second paragraphs of **Study variables** for the outcome, explanatory and confounding variables
    -   third paragraph of the **Statistical analyses** for the survey weights variable.

    2.  There were **five** potential confounders.
    3.  Potentially useful functions for this exercise:

    -   [%in%](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/match)
    -   [levels](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/levels)
    -   [recode](https://rdrr.io/cran/car/man/recode.html)
    -   [subset](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/subset)
    -   [as.factor](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/factor)
    -   [relevel](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/relevel)
    -   or [dplyr ways](https://ehsanx.github.io/intro2R/data-analysis-with-dplyr.html): [filter](https://www.rdocumentation.org/packages/dplyr/versions/0.7.8/topics/filter), [select](https://www.rdocumentation.org/packages/dplyr/versions/0.7.8/topics/select)

eight variables (one outcome, one exposure, five confounders, and one survey weight)
outcome: SCR_082
exposure: GEN_10
confounders: DHHGAGE
weight: WTS_M


The outcome variable was categorized into three groups: excellent or very good, good, fair or poor.

age (grouped in 10-year increments), 
sex (male versus female), 
respondent’s cultural or racial identity (white versus non-white), 
main source of income (employment income, workers compensation/welfare, senior’s benefits, other, not applicable and not stated), 
having received help for problems with emotions, mental health or the use of alcohol/drugs in the previous 12 months (yes versus no).

```{r}
# your code here
table(cmh_sub$SCR_082)
table(cmh_sub$GEN_10)
library(dplyr)
NAs <- c("NOT APPLICABLE", "DON'T KNOW", "REFUSAL", "NOT STATED")
cmh_sub_tidy <- dplyr::select(cmh_sub, ADM_RNO, SCR_082, GEN_10, DHHGAGE, DHH_SEX, SDCGCGT, INCG7, PNC_01A, WTS_M) %>% 
  dplyr::mutate(outcome=dplyr::case_when(SCR_082 %in% c("EXCELLENT", "VERY GOOD") ~ "excellent",
                                  SCR_082 == "GOOD" ~ "good",
                                  SCR_082 %in% c("FAIR", "POOR") ~ "poor"),
                income=if_else(INCG7 %in% c("DON'T KNOW", "REFUSAL", "NOT STATED"), NA,  INCG7),
                age=case_when(DHHGAGE %in% c("15 TO 19 YEARS", "20 TO 24 YEARS") ~"15-24",
                              DHHGAGE %in% c("25 TO 29 YEARS", "30 TO 34 YEARS") ~"25-34",
                              DHHGAGE %in% c("35 TO 39 YEARS", "40 TO 44 YEARS") ~"35-44",
                              DHHGAGE %in% c("45 TO 49 YEARS", "50 TO 54 YEARS") ~"45-54",
                              DHHGAGE %in% c("55 TO 59 YEARS", "60 TO 64 YEARS") ~"55-64",
                              DHHGAGE %in% c("65 TO 69 YEARS", "70 TO 74 YEARS", "75 TO 79 YEARS", "80 YEARS OR MORE") ~">=65"),
                sex=if_else(DHH_SEX %in% NAs, NA, DHH_SEX ),
                Community_belonging=if_else(GEN_10 %in% NAs, NA, GEN_10 ),
                Race=if_else(SDCGCGT %in% NAs, NA, SDCGCGT )
                ) 

cmh_sub_tidy_selected <- cmh_sub_tidy[,c("outcome","Community_belonging", "sex","Race", "age", "income")]

cmh_sub_tidy_selected$outcome <- factor(cmh_sub_tidy_selected$outcome, c("poor", "good", "excellent"))
cmh_sub_tidy_selected$Community_belonging <- factor(cmh_sub_tidy_selected$Community_belonging, 
                                                    c("VERY WEAK", "SOMEWHAT WEAK", "SOMEWHAT STRONG", "VERY STRONG"))
cmh_sub_tidy_selected$sex <- factor(cmh_sub_tidy_selected$sex, c("MALE", "FEMALE"))
cmh_sub_tidy_selected$age <- factor(cmh_sub_tidy_selected$age, c("15-24", "25-34", "35-44", "45-54", "55-64", ">=65" ))
cmh_sub_tidy_selected$income <- factor(cmh_sub_tidy_selected$income, c("EMPLOYMENT INC.", "EI/WORKER'S COMP", "SENIOR BENEFITS", "OTHER"))
cmh_sub_tidy_selected$Race <- factor(cmh_sub_tidy_selected$Race, c("WHITE", "NON-WHITE"))

cmh_sub_tidy_selected <- na.omit(cmh_sub_tidy_selected)
tab_CCHL <- tableone::CreateTableOne(data=cmh_sub_tidy_selected, vars=c("Community_belonging", "sex","Race", "age", "income"), strata=c("outcome"), includeNA = TRUE)
print(tab_CCHL)
```

### 1(d) Creating analytic dataset

Outcome variable has a category 'NOT STATED', but for our analysis, we will omit anyone associated with this category. Similarly, for explanatory variable, we have categories such as DON'T KNOW, REFUSAL and NOT STATED. We will omit anyone with these categories.

-   Assign missing values for categories such as DON'T KNOW, REFUSAL and NOT STATED.
-   Recode the variables as shown in Table 1 in the article. You can use any function/package of your choice. Here is an example (but feel free to use other functions. In R there are many other ways to do this same task.

```{r}
## your code here
# levels(your.data.frame$your.age.variable) <- 
#   list("15 to 24 years" = c("15 TO 19 YEARS", "20 TO 24 YEARS"),
#        "25 to 34 years" = c("25 TO 29 YEARS", "30 TO 34 YEARS"),
#        "35 to 44 years" = c("35 TO 39 YEARS", "40 TO 44 YEARS"),
#        "45 to 54 years" = c("45 TO 49 YEARS", "50 TO 54 YEARS"),
#        "55 to 64 years" = c("55 TO 59 YEARS", "60 TO 64 YEARS"),
#        "65 years or older" = c("65 TO 69 YEARS", "70 TO 74 YEARS", 
#        "75 TO 79 YEARS", "80 YEARS OR MORE"))
```

### 1(e) Number of columns and variable names

Report the number of columns in your analytic dataset, and the variable names.

```{r}
# your code here
colnames(cmh_sub_tidy)
```

## Question 2: Table 1 \[20% grade\]

Reproduce Table 1 presented in the article (or see below). Omit the 'Main source of income' variable from the table. The table you produce should report numbers as follows, with all columns as shown in the table. In other words, the numbers should match.

| Self-rated Mental Health Variable | Total n(%)  | Poor or Fair n(%) | Good n(%)   | Very good or excellent n(%) |
|---------------|---------------|---------------|---------------|---------------|
| **Study sample**                  | 2628 (100)  | 1002 (38.1)       | 885 (33.7)  | 741 (28.2)                  |
| **Community belonging**           |             |                   |             |                             |
| \- Very weak                      | 480 (18.3)  | 282 (28.1)        | 118 (13.3)a | 80 (10.8)a                  |
| \- Somewhat weak                  | 857 (32.6)  | 358 (35.7)        | 309 (34.9)  | 190 (25.6)                  |
| \- Somewhat strong                | 1005 (38.2) | 288 (28.7)        | 362 (40.9)  | 355 (47.9)                  |
| \- Very strong                    | 286 (10.9)  | 74 (7.4)a         | 96 (10.8)a  | 116 (15.7)a                 |
| **Sex**                           |             |                   |             |                             |
| \- Females                        | 1407 (53.5) | 616 (61.5)        | 487 (55.0)  | 304 (41.0)                  |
| \- Males                          | 1221 (46.5) | 386 (38.5)        | 398 (45.0)  | 437 (59.0)                  |
| **Age group**                     |             |                   |             |                             |
| \- 15 to 24 years                 | 740 (28.2)  | 191 (19.1)        | 264 (29.8)  | 285 (38.5)                  |
| \- 25 to 34 years                 | 475 (18.1)  | 141 (14.1)        | 167 (18.9)  | 167 (22.5)                  |
| \- 35 to 44 years                 | 393 (15.0)  | 185 (18.5)        | 119 (13.4)a | 89 (12.0)a                  |
| \- 45 to 54 years                 | 438 (16.6)  | 220 (22.0)        | 139 (15.7)  | 79 (10.7)a                  |
| \- 55 to 64 years                 | 379 (14.4)  | 198 (19.7)        | 113 (12.8)a | 68 (9.2)a                   |
| \- 65 years or older              | 203 (7.7)   | 67 (6.6)a         | 83 (8.4)a   | 53 (7.1)b                   |
| **Race/Ethnicity**                |             |                   |             |                             |
| \- Non-white                      | 458 (17.4)  | 184 (18.4)        | 140 (15.8)  | 134 (18.1)                  |
| \- White                          | 2170 (82.6) | 818 (81.6)        | 745 (84.2)  | 607 (81.9)                  |
| **Main source of income**         |             |                   |             |                             |
| \- Employment Income\^d           | 1054 (40.1) | 289 (28.8)        | 386 (43.6)  | 379 (51.1)                  |
| \- Worker's Compensation\^e       | 160 (6.1)   | 91 (9.1)a         | 44 (5.0)b   | 25 (3.4)c                   |
| \- Senior Benefits\^f             | 134 (5.1)   | 57 (5.7)a         | 42 (4.7)b   | 35 (4.7)                    |
| \- Other\^g                       | 184 (7.0)   | 82 (8.2)a         | 60 (6.8)a   | 42 (5.7)b                   |
| \- Not applicable\^h              | 851 (32.4)  | 402 (40.1)        | 263 (29.7)  | 186 (25.1)                  |
| \- Not Stated\^i                  | 245 (9.3)   | 81 (8.1)a         | 90 (10.2)a  | 74 (10.0)                   |

$^a$ Coefficient of variation between 16.6 and 25.0%. $^b$ Coefficient of variation between 25.1 and 33.3%. $^c$ Coefficient of variation \> 33.3%. $^d$ Employment Income: Wages/salaries or self-employment. $^e$ Worker's compensation: Employment insurance or worker's compensation or social assistance/welfare. $^f$ Senior Benefits: Benefits from Canada or Quebec Pension Plan or job related retirement pensions, superannuation and annuities or RRSP/RRIF of Old Age Security and Guaranteed Income Supplement. $^g$ Other: Dividends/interest or child tax benefit or child support or alimony or other or no income. $^h$ Not applicable: Respondents who live in a household with only one person. The income variable "main source of personal income" is applicable only to those that live in a household of more than one person. $^i$ Not Stated: Question was not answered (don't know, refusal, not stated).

```{r,warning=FALSE,message=FALSE}
# your code here
require(tableone)

```

## Question 3: \[20% grade\]

### 3(a) Subset

Subset the dataset excluding 'Very good or excellent' responses from the self-rated mental health variable

```{r}
# your code here
cmh_sub1 <- dplyr::filter(cmh_sub_tidy_selected, outcome != "excellent")
cmh_sub1$outcome <- factor(cmh_sub1$outcome, c("poor","good"))
```

### 3(b) Recode

Recode self-rated mental health variable and make it a binary variable: 'Good' vs. 'Poor' (simplifying category labels only). Convert that variable to a factor variable with 'Poor' being the reference level.

```{r}
# your code here
cmh_sub1
fit1 <- glm(I(outcome == "good") ~ Community_belonging +age + sex + Race, data = cmh_sub1, family = binomial)
Publish::publish(fit1)
```

### 3(c) Regression

Run a logistic regression model for finding the relationship between community belonging (Reference: Very weak) and self-rated mental health (Reference: Poor) among respondents with mental or substance use disorders. Adjust the model for three confounders: sex, age, and race/ethnicity. Do not need to report summary of the model.

```{r}
# your code here

```

### 3(d) Reporting odds ratio

Report the odds ratios and associated confidence intervals. Publish or jtools package could be useful to report the odds ratios with confidence intervals.

```{r,warning=FALSE,message=FALSE}
# your code here

```
